---
title: "WikidataQueryServiceR"
output:
  github_document:
    toc: true
    toc_depth: 3
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
# install.packages("printr", type = "source", repos = c("https://yihui.name/xran", CRAN = "https://cran.rstudio.com"))
library(printr)
```

[![Project Status: Active – The project has reached a stable, usable state and is being actively developed.](http://www.repostatus.org/badges/latest/active.svg)](http://www.repostatus.org/#active)
[![CRAN_Status_Badge](http://www.r-pkg.org/badges/version/WikidataQueryServiceR)](https://cran.r-project.org/package=WikidataQueryServiceR)
[![CRAN Total Downloads](https://cranlogs.r-pkg.org/badges/grand-total/WikidataQueryServiceR)](https://cran.r-project.org/package=WikidataQueryServiceR)
[![License: MIT](https://img.shields.io/badge/license-MIT-blue.svg)](https://opensource.org/licenses/MIT)

This is an R wrapper for the [Wikidata Query Service (WDQS)](https://www.mediawiki.org/wiki/Wikidata_query_service) which provides a way for tools to query [Wikidata](https://www.wikidata.org/wiki/Wikidata:Main_Page) via [SPARQL](https://en.wikipedia.org/wiki/SPARQL) (see the beta at https://query.wikidata.org/). It is written in and for R, and was inspired by Oliver Keyes' [WikipediR](https://github.com/Ironholds/WikipediR) and [WikidataR](https://github.com/Ironholds/WikidataR) packages.

__Author:__ Mikhail Popov (Wikimedia Foundation)<br/> 
__License:__ [MIT](http://opensource.org/licenses/MIT)<br/>
__Status:__ Active

## Installation

```R
install.packages("WikidataQueryServiceR")
```
    
To install the development version:

```R
# install.packages(c("devtools", "httr", "dplyr", "jsonlite"))
devtools::install_github("bearloga/WikidataQueryServiceR")
```

## Usage

```{r load}
library(WikidataQueryServiceR)
```

You submit SPARQL queries using the `query_wikidata()` function.

### Example: fetching genres of a particular movie

In this example, we find an "instance of" ([P31](https://www.wikidata.org/wiki/Property:P31)) "film" ([Q11424](https://www.wikidata.org/wiki/Q11424)) that has the label "The Cabin in the Woods" ([Q45394](https://www.wikidata.org/wiki/Q45394)), get its genres ([P136](https://www.wikidata.org/wiki/Property:P136)), and then use [WDQS label service](https://www.mediawiki.org/wiki/Wikidata_query_service/User_Manual#Label_service) to return the genre labels.

```{r wdqs_example, cache=TRUE}
query_wikidata('SELECT DISTINCT
  ?genre ?genreLabel
WHERE {
  ?film wdt:P31 wd:Q11424.
  ?film rdfs:label "The Cabin in the Woods"@en.
  ?film wdt:P136 ?genre.
  SERVICE wikibase:label { bd:serviceParam wikibase:language "en". }
}')
```

For more example SPARQL queries, see [this page](https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/queries/examples) on [Wikidata](https://www.wikidata.org/wiki/Wikidata:Main_Page).

`query_wikidata()` can accept multiple queries, returning a (potentially named) list of data frames. If the vector of SPARQL queries is named, the results will inherit those names.

### Fetching queries from Wikidata's examples page

The package provides a [WikipediR](https://github.com/Ironholds/WikipediR/)-based function for getting SPARQL queries from the [WDQS examples page](https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/queries/examples).

```{r get_examples, cache=TRUE}
sparql_query <- get_example(c("Cats", "Horses", "Largest cities with female mayor"))
```
```{r, eval=FALSE}
sparql_query[["Largest cities with female mayor"]]
```
```{r, echo=FALSE, results='asis'}
cat("```SPARQL\n", sparql_query[["Largest cities with female mayor"]], "\n```")
```

Now we can run all three extracted SPARQL queries and get back three data.frames:

```{r run_examples, cache=TRUE, dependson='get_examples'}
results <- query_wikidata(sparql_query)
results$`Largest cities with female mayor`[, c("cityLabel", "mayorLabel")]
```

## Links for learning SPARQL

- [A beginner-friendly course for SPARQL](https://www.wikidata.org/wiki/Wikidata:A_beginner-friendly_course_for_SPARQL)
- Building a SPARQL query: [Museums on Instagram](https://www.wikidata.org/wiki/Help:SPARQL/Building_a_query/Museums_on_Instagram)
- [SPARQL Query Examples](https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/queries/examples) for WDQS
- [Using SPARQL to access Linked Open Data](http://programminghistorian.org/lessons/graph-databases-and-SPARQL) by Matthew Lincoln
- Interesting or illustrative [SPARQL queries](https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/queries) for Wikidata
- Wikidata [2016 SPARQL Workshop](https://www.wikidata.org/wiki/Wikidata:SPARQL_query_service/2016_SPARQL_Workshop)
- [Wikidata SPARQL Query video tutorial](https://www.youtube.com/watch?v=1jHoUkj_mKw) by Navino Evans
- _[Learning SPARQL](http://www.learningsparql.com/)_ by Bob DuCharme
- [WDQS User Manual](https://www.mediawiki.org/wiki/Wikidata_query_service/User_Manual)

## Additional Information

Please note that this project is released with a [Contributor Code of Conduct](https://github.com/bearloga/WikidataQueryServiceR/blob/master/CONDUCT.md). By participating in this project you agree to abide by its terms.
